---
title: "HW5"
author: "Brandon Xicon:xiconb"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*Note that reading data and loading packages is hidden on the HTML for readabilty*

## Liquor Sales in Iowa

This dataset https://data.iowa.gov/Sales-Distribution/2019-Iowa-Liquor-Sales/38x4-vs5h contains the spirits purchase information of Iowa Class “E” liquor licensees by product and date of purchase for 2019. The dataset can be used to analyze total spirits sales in Iowa of individual products at the store level.

**Submission instructions**: Create a folder named `ds202_hw5`, and name the RMarkDown file including your solutions `hw5.Rmd` under this folder. For the submission, create a GitHub repository named `ds202_hw5` under your GitHub account, and push both `hw5.Rmd` and the knitted `hw5.html` before the deadline. I will assume you use the same GitHub username as for your HW4 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on April 10** by 11:59 pm.

1. The complete dataset is too large. Download and import the alcohol sales data for Ames only: 

  a. Go to https://data.iowa.gov/Sales-Distribution/2019-Iowa-Liquor-Sales/38x4-vs5h
  b. Click Filter -- Filter -- Go advanced now
  c. Add a New Filter Condition: City is Ames.
  d. Export -- CSV -- Download
  e. Use `readr::read_csv` to import.
```{r include=FALSE}
library(dplyr)
library(lubridate)
library(tidyverse)
Liquor<-readr::read_csv('2019_Iowa_Liquor_Sales_Ames.csv')
Liquor
```
2. Data cleaning: 
  a. extract geographic latitude and longitude from the variable `Store Location`
  b. make sure the variables are in the correct type, by converting incorrect variable types.
  c. extract year, month and day from the variable `Date` as separate columns.
```{r}
#Getting longitude and latitude from Store Location
Liquor <- Liquor %>%
  na.omit %>%
  mutate(`Store Location`=substr(`Store Location`, 8, nchar(`Store Location`) - 1)) %>%
  separate(col='Store Location', into=c('Latitude', 'Longitude'), sep=" ")

#changing values to numeric

Liquor<-Liquor %>%  
  mutate(Latitude = as.numeric(Latitude)) %>%
  mutate(Longitude = as.numeric(Longitude)) %>%
  mutate(`Vendor Number` = as.numeric(`Vendor Number`))

#Getting day month year from date as seperate columns

Liquor <- Liquor %>% 
  separate('Date', c('Month', 'Day', 'Year'), sep='/') %>%
  mutate(Month = as.numeric(Month)) %>%
  mutate(Day = as.numeric(Day)) %>%
  mutate(Year = as.numeric(Year))
Liquor  
```
3. First overview:
  a. Plot a scatterplot of lat and long of store locations. 
```{r}
Liquor %>% ggplot(aes(x=Longitude, y=Latitude)) + geom_point() + 
  ggtitle("Latitude vs Longitude of Liquor Stores in Ames")
```
  b. Provide a visual breakdown of the liquor category (by `Category Name`). Include volume sold in the breakdown.
```{r}
#ggplot(Liquor,aes(x=`Category Name`, count=`Volume Sold (Liters)`)) + 
 # geom_bar(position=position_dodge(width = .5)) + 
  #ylab('Volume Sold (Liters)') + 
  #ggtitle("Voume of Alcohol Sold in Each Category")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
Liquor %>%
  filter(!is.na(`Category Name`)) %>%
  group_by(`Category Name`) %>%
  summarize(`Volume(Liters)` = sum(`Volume Sold (Liters)`)) %>%
  ggplot(aes(x = `Category Name`, y = `Volume(Liters)`)) +
  geom_bar(stat = "identity") +
ggtitle("Voume of Alcohol Sold in Each Category")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```
4. Find the daily sales liquor sales in Ames in 2019. Summarize the number of sales, volume of liquor sold, and amount of money spent.
```{r}
Liquor2<-Liquor %>%
  group_by(Month, Day,Year) %>%
  summarize(`Number of Sales` = n(), `Total Volume Sold (Liters)` = sum(`Volume Sold (Liters)`),
            `Money Spent` = sum(`Sale (Dollars)`)) %>%
  arrange(Month,Day,Year,desc(`Number of Sales`))
Liquor2
```
5. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Discover trends with the help of `geom_smooth`. Describe any patterns.
```{r}
ggplot(Liquor2,aes(x=Day,y=`Total Volume Sold (Liters)`))+facet_wrap(. ~Month)+geom_point()+geom_smooth()+
  ylim(0, 15000)
```
6. Find the dates for ISU football home games in Fall 2019. Can you see a pattern? 
`
home games were 8/31,9/14,9/21,10/5,10/26,11/16,11/23
These points do not exist in our dataframe however based on the best fit lines above we see that the spikes or high points in each month seemed to correspond to the home games. 

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio. 
